master thesis

### 2018-03-20 Paper - Latex/Paper - Skizze - v4.tex
Paper - Introduction

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Introduction}
  This paper is about the analysis of ancestral state reconstruction algorithms for non-binary trees, 
    applied to the currently largest phylogeny synthesis tree of Open Tree Of Life, with the 
    application of prediciton of parasitism. \\

  \anmerkung{Mein Vorschlag einer Gliederung (jeweils ca. ein Absatz) (Bernhard)}
  \begin{enumerate}
    \item Motivation:
      \begin{itemize}
        \item Was ist das große Ziel? \\
          Das Ziel dieser Arbeit ist die Anwendung von maximum parsiomony algorithmen auf nicht 
          binäre Bäume und auf sehr große Datensätze. Insbesondere auf das Beispiel 'Entstehung des 
          Parasitismus' im ganzen Eukaryotischen Tree of Life.
        \item Was soll erreicht werden? \\
          Wir wollen vorhandene Algorithm (Sankoff/castor \cite{Louca2017}) auf diese Aufgabenstellung hin testen 
          und ihre Vorhersagekraft abschätzen. Außerdem wollen wir den Fitch algorithmus für binäre 
          Bäume auf unser Problem erweitern und mit dem Sankoff Algorithmus vergleichen.
        \item Warum ist das relevant? Was könnte man dann tun?\\
          predict states of species... \\
          \todo{!!!}
      \end{itemize}
    \item Hintergrund:
      \begin{itemize}
        \item Was gab es in dieser Richtung bereits als ganze Ansätze oder wenn nicht, warum nicht?
          Woran ist es bisher gescheitert? \\
          For about 50 years, people have been working on ancestral state reconstruction. The first paper 
            mentioned above is by Camin and Sokal, who in 1965 were working on algorithms for discrete-state 
            data \cite{Camin1965}. Different methods have been developed and the question is which method is 
            the most suitable for our problem. \\
          For this purpose, various studied methods and their advantages and disadvantages were compared. \\

          Bisher wurden vorallem Algorithmen für das binäre Problem entwickelt, da man wesentlich 
            kleinere Teilbäume betrachtet hat, von welchen man auch alle Aufspaltungen kennt. Durch 
            die Entwicklung von OTL, eines gesamten Baum des Lebens, ergibt sich das Problem, dass 
            dieser bei weitem nicht binär ist. \\ \\
  Researchers of the phylogenies have been dealt with the ancenstral state reconstruction in the 
    60s. The first methods were only brute force \todo{Quelle, siehe Fitch: Camin and Sokal 1965}. 
    Next came a set of parsimony algorithms such as: Fitch-parsimony \cite{Fitch1971}, 
    Wagner-parsimony \cite{Swofford1987} ... \todo{weitere?}. \\
    With more and more data, there is now the possibility to use more information to calculate the 
    probabilities of the ancestral states. In addition to the states of the leafs, algorithms could 
    also use branch lengths. The likelihood based algorithms came more in interest. \\
    Our focus came with another 'data extension'. We wanted to work with the biggest phylogenetic tree 
    that exists at this moment, which goes over all observed species. For most \todo{most?} species 
    there is no phylogeny, but only a taxonomic classification.

        \item Welche Grundlagen sind notwendig:
          \begin{itemize}
            \item open tree of life: Was ist das, warum relevant und überlegen als reine Ansätze? \\
              \todo{!!!} \\ \\
  So the biggest 'phylogenetic tree' is a synthesis 
    of phylogenetic trees filled with a taxonomic tree given by Open Tree of Life \cite{Hinchliff2015}.
    This tree is not binary and therefore the developed algorithms are not directly applicable. \\
            \item Algorithmen: Was gibt es? Ruhig ausführlicher als hier bereits und vor allem auch 
              nach einer Darstellung am Ende ableiten, was für uns relevant ist. Also beschreiben, 
              wie Methode a, b, c funktionieren und dann abwägen, was daher für Dich am 
              relevantesten ist. \\
              \todo{!!!} \\
              \anmerkung{GloBI und OTL in der Einleitung vorstellen. (Emanuel)}
          \end{itemize}
      \end{itemize}
    \item Outlook/Structure of this work
  \end{enumerate}

  In this work, we have looked at the algorithms that are generally suited to our data, to develop 
  them further for the not binary case, and finally to compare their usability with our sythesis 
  tree. \\
  We have decided to consider only parsimony algorithms since we have no information on branch 
  lengths and no other additional information like different transition probabilities of our states.

  In den Bacteria und Archaea wurde das mal allgemein für binary traits gemacht:
  "A total 90~\% of all binary traits described molecular functions, specifically the presence or absence of a gene or a set of gene involved in biochemical pathways."
  \cite{Goberna2015} \\
  Unterschied bei uns ist parasitismus komplexer.. da sind meist nicht einzelne Gene zuständig..

>>>>

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Introduction}
  This paper is about the analysis of ancestral state reconstruction algorithms for non-binary trees, 
    applied to the currently largest phylogeny synthesis tree of Open Tree Of Life, with the 
    application of prediciton of parasitism. \\
  \todo{Or:} The aim of this thesis is the application of maximum parsiomony algorithms to non-binary 
    trees and very large datasets. In particular, the example 'Origins of Parasitism' throughout the 
    Eukaryota Tree of Life. \\
  We will test the existing algorithms Sankoff (R package \textit{Castor}\cite{Louca2017}) and Fitch 
    for this task and estimate their predictive power. In addition, we will extend the Fitch 
    algorithm from binary trees to multifurcating trees to make it usable and comparable for our 
    current problem. \\
  This achieves that we can predict ancestral states and unknown states of living species for large 
    non-binary relatives trees. \\

  For about 50 years, people have been working on ancestral state reconstruction. The first paper 
    mentioned above is by Camin and Sokal, who in 1965 were working on algorithms for discrete-state 
    data \cite{Camin1965}. Different methods have been developed and the question is which method is 
    the most suitable for our problem. \\

  \todotext{In den Bacteria und Archaea wurde das mal allgemein für binary traits gemacht: \\
  "A total 90~\% of all binary traits described molecular functions, specifically the presence or absence of a gene or a set of gene involved in biochemical pathways."
  \cite{Goberna2015} \\
  Unterschied bei uns ist parasitismus komplexer.. da sind meist nicht einzelne Gene zuständig..} \\

  For this purpose, various studied methods and their advantages and disadvantages will be compared. \\
  These algorithms were developed especially for the binary problem, since one has considered much 
    smaller subtrees, of which one knows all the splits. \\
  By developing a whole of an entire tree of life, the problem arises that it is by no means binary.

  % Researchers of the phylogenies have been dealt with the ancenstral state reconstruction in the 
  %   60s. The first methods were only brute force \todo{Quelle, siehe Fitch: Camin and Sokal 1965}. 
  %   Next came a set of parsimony algorithms such as: Fitch-parsimony \cite{Fitch1971}, 
  %   Wagner-parsimony \cite{Swofford1987} ... \todo{weitere?}. \\
  %   With more and more data, there is now the possibility to use more information to calculate the 
  %   probabilities of the ancestral states. In addition to the states of the leafs, algorithms could 
  %   also use branch lengths. The likelihood based algorithms came more in interest. \\
  %   Our focus came with another 'data extension'. We wanted to work with the biggest phylogenetic tree 
  %   that exists at this moment, which goes over all observed species. For most \todo{most?} species 
  %   there is no phylogeny, but only a taxonomic classification.
  To accomplish this task, a large phylogenetic tree and information about the current species states 
    is needed. %In this work we use the synthesis tree of Open Tree of Life (OTL) \cite{Hinchliff2015} 
  The biggest 'phylogenetic tree' is a synthesis of phylogenetic trees filled with a taxonomic tree 
    given by Open Tree of Life \cite{Hinchliff2015}. This tree is not binary and therefore the 
    developed algorithms are not directly applicable.Therefore, an algorithm must be found that can 
    work with non-binary trees. \\
  For the information about the current states of the species we use the interaction database Global 
    Biotic Interactions (GloBI) \cite{Poelen2014}.

  % In this work, we have looked at the algorithms that are generally suited to our data, to develop 
  % them further for the not binary case, and finally to compare their usability with our sythesis 
  % tree. \\
  % We have decided to consider only parsimony algorithms since we have no information on branch 
  % lengths and no other additional information like different transition probabilities of our states.
